The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-12-10 Manuscript_draft/Draft.Rmd
small things

# Introduction
The soil is one of the largest terrestrial reservoirs of carbon (C), containing approximately three times more C than the atmosphere or the vegetation on Earth (IPCC, 2021). Soil C is released to the atmosphere as carbon dioxide (CO2) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration). This process (i.e., soil respiration) constitutes the main C efflux into the atmosphere (ref). Under the new climate change scenarios increases of CO2 efflux from soils are expected (Schlesinger and Andrews, 2000), this can imply a positive feedback to future warming if stored C is mobilized and C input to soil does not increase (Lamberty&Thomson, 2010). Because soil respiration is highly sensitive to soil temperature and soil moisture (Nissan et al. 2022), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management, is critical for predicting future atmospheric CO2 concentrations. Temperature affects respiration rates positively (Lloyd and Taylor, 1994), by increasing the metabolic rates of soil microorganisms and enzyme efficiency (ref). However, the relationship with soil moisture is non-linear. Low water content limit nutrient diffusion and enzymes and exert drought-stress on microbes while high water contents limits oxygen (O2) diffusion, creating anaerobic conditions that suppress aerobic microbial respiration (ref). Something here about modelling?

>>>>

# Introduction
The soil is one of the largest terrestrial reservoirs of carbon (C), containing approximately three times more C than the atmosphere or the vegetation on Earth (IPCC, 2021). Soil C is released to the atmosphere as carbon dioxide (CO2) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration). This process (i.e., soil respiration) constitutes the main C efflux into the atmosphere (ref). Under the new climate change scenarios increases of CO2 efflux from soils are expected (Schlesinger and Andrews, 2000), this can imply a positive feedback to future warming if stored C is mobilized and C input to soil does not increase (Lamberty&Thomson, 2010). Because soil respiration is highly sensitive to soil temperature and soil moisture (Nissan et al. 2022), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest disturbances, is critical for predicting future atmospheric CO2 concentrations. Temperature affects respiration rates positively (Lloyd and Taylor, 1994), by increasing the metabolic rates of soil microorganisms and enzyme efficiency (ref). However, the relationship with soil moisture is non-linear. Low water content limit nutrient diffusion and enzymes and exert drought-stress on microbes while high water contents limits oxygen (O2) diffusion, creating anaerobic conditions that suppress aerobic microbial respiration (ref). Something here about modelling?
